### 2017-06-07 paper/abstract.tex
abstract added

In Biology, there are several questions that translate to combinatorial 
search.
%
For example, vesicle traffic systems in biological cells exhibit
several graph properties such as three connectivity and
a natural question arises for biologists that what are all possible
networks for various combinations of those properties.
%
It will help biologists for .....
%
In this paper, we present a SAT/QBF encodings of the properties over
vesicle traffic systems and a tool that searches for the networks
that satisfies the properties using SAT/QBF solvers.
%
In our experiments, we show that our tool can search for networks of
sizes that are considered to be relevant by biologists.

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: