### 2017-06-12 paper/main.tex
some reorg

\documentclass{llncs}

\input{macro}

\begin{document}

\title{SAT Solving for Vesicle Traffic Systems in Cells}

\author{to be announced}

\date{\today}

\maketitle

% \vspace*{0.2in}
% % Title must be 250 characters or less.
% \begin{flushleft}
% {\Large
% \textbf\newline{ Vesicle traffic system and constraints.} % Please use 
% }
% \newline
% \\
% \end{flushleft}

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Biological Problem}
\label{sec:bio}
% \input{bio}

\section{The Problem Encoding}
\label{sec:encoding}
\input{encoding}

\section{Implementation and Experiments}
\label{sec:experiments}
\input{experiments}

\section{Related Work}
\label{sec:related}
\input{related}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: