### 2017-06-16 paper/conclusion.tex
added

\ankit{interpret experiments (3-4 lines)}

\srivas{Future work(quantifier problem etc)}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: