### 2017-06-16 paper/abstract.tex
second half of the intro

In biology, there are several questions that translate to combinatorial
search. For example, vesicle traffic systems that move cargo within
eukaryotic cells have been proposed to exhibit several graph properties
such as three connectivity. These properties are consequences of underlying
biophysical constraints. A natural question for biologists is: what are all
possible networks for various combinations of those properties? In this
paper, we present novel SAT encodings of the properties over vesicle traffic
systems and a tool that searches for the networks that satisfies the
properties using SAT solvers. In our experiments, we show that our tool
can search for networks of sizes that are considered to be relevant by
biologists.

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: