### 2017-06-20 paper/bio.tex

Again about two pages long

\begin{itemize}
\item \mukund{Detailed BIO}
\item \srivas{Detailed BIO to Graph problem}
\end{itemize}

The following terms must be defined in this section

\begin{itemize}
\item Vesicular traffic network.
\item Recycling of molecules, steady state.
\item Maximum connectivity (rather than necc and suff cond) and minimal connective required by the graph. 
\item Connected-ness. 3 and 4.
\item Fusion rules, inhibition/regulation mechanism, pairing mechanism and matrix.
\item Property of interest.
\item Reachability definition (I defer).
\end{itemize}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: