### 2017-07-02 paper/bio.tex
added-bio-section2

Again about two pages long

\begin{itemize}
\item \mukund{Detailed BIO}
The cell as a transport graph: We consider a cell to be a collection of
compartments (nodes) and vesicles (edges), thus defining a transport graph.
Compartments represent the large membrane-bound organelles, which are
connected by constant fluxes of small transport vesicles. Every compartment
or vesicle has a set of molecules associated with it. These include SNAREs
and other biomolecules that regulate membrane traffic, or that play the
role of passive cargo. Here we consider an abstract set of molecules and do
not explicitly associate them with particular protein varieties.

Molecular flows and steady state: The total amount of each molecular type
on each compartment can increase or decrease: this is due to gain or loss
of that molecular type driven by incoming or outgoing vesicles. Each edge
is thus associated with a flux of all the molecular types carried by the
corresponding vesicle. We assume the cell is in a steady state where each
compartments composition does not vary over time, meaning that all
incoming and outgoing fluxes are balanced for each molecular type at each
compartment.

Vesicle targeting driven by molecular interactions: The model specification
includes a description of how molecular properties influence vesicle
transport. A vesicle can only contain a set of molecules that are present
on its source compartment. Once a vesicle has budded out of the source, the
molecules it carries determine its properties. In particular, for any given
pair of a vesicle and a compartment, the set of molecules that label the
former and latter determine whether the vesicle will fuse to that
compartment. Biophysically, fusion requires a direct physical interaction
between at least one molecular type on the vesicle and one molecular type
on the compartment. The list of molecular pairs that can drive a fusion
event is given an a fusion pairing matrix.

Molecular regulation: The final layer of our framework involves how the
molecules are regulated. We assume that for fusion to occur, the pair of
molecular types involved on the vesicle and compartment must both be in an
“active” state. Whether these molecules are active or inactive depends on
the remaining molecules found on the vesicle or compartment, respectively.
This embodies the biological fact that other molecules can regulate the
fusion-driving molecules. We test many different versions of molecular
regulation. Most generally, the activity state of a given molecule can be a
Boolean function of all the molecular types on a compartment or vesicles.
We have also tested a particularly simple regulation mechanism in which two
molecules that can pair to drive fusion inhibit one another. This is
motivated by the idea that pairing must generate an inactive bi-molecular
complex.

Synthesis: We now combine all the above ingredients. Given a particular
transport graph, a particular labeling of all the compartments and edges,
a particular fusion pairing matrix, and a particular regulatory model we do
the following. 1. We determine which molecules are “active” on every
compartment or vesicle. 2. For every vesicle fusing to a compartment, we
determine whether there exists an active pair (one molecule on the vesicle,
one on the compartment) which drives that fusion event. 3. For every
vesicle-compartment pair where the vesicle does not fuse to the
compartment, we verify that there is no pairing of active molecules on the
vesicle and compartment that could drive their fusion. 4. We verify that
every molecular type entering a compartment also leaves the compartment,
and also that every molecular type entering a set of compartments also
leaves that set; this is the steady state condition. This tells us that the
particular graph and molecular labeling does represent an allowed steady
state configuration of a VTS.

\item \srivas{Detailed BIO to Graph problem}
\end{itemize}

The whole cell should be in stable condition i.e flow of the molecule
is balanced at every point and globally. Each molecule leaving its
source node on a vesicle should come back in a cycle.

The following terms must be defined in this section

\begin{itemize}
\item Vesicular traffic network.
\item Recycling of molecules, steady state.
\item Maximum connectivity [LGC, Least guarantee connectivity] and minimal connective [LRC (least required connectivity)] (rather than necc and suff cond) required by the graph. 
\item Connected-ness. 3 and 4.
\item Pairing mechanism (matrix).
\item Fusion rules, inhibition/regulation mechanism, Pairing inhibition.
\item Potential to connect.
\item Property of interest.
\item Reachability definition.
\end{itemize}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: