### 2017-07-11 paper/pm-1-matrix.tex
added dia

\begin{figure}
\centering
\begin{tikzpicture}
\matrix [matrix of math nodes,left delimiter=(,right delimiter=),row sep=0.5cm,column sep=0.5cm] (m) {
\times & \times & \times &\times&\times&\times & \times & \times\\
\times & \times & \times&\times&\times&\times& 1 &\times \\
\times & \times & \times&\times&\times&1&\times &\times\\
\times & \times & \times&\times & 1 & \times &\times & \times\\
 \times & \times & \times & \times & \times & \times & \times & \times\\
 \times & \times & \times &\times & \times & \times & \times & \times\\
 \times & \times & \times & \times & \times & \times & \times & \times\\
 \times & \times & \times & \times & \times & \times&\times & \times\\};

\draw[dashed] ($0.5*(m-1-4.north east)+0.5*(m-1-5.north west)$) --
     ($0.5*(m-8-5.south east)+0.5*(m-8-4.south west)$);

\draw[dashed] ($0.5*(m-4-1.south west)+0.5*(m-5-1.north west)$) --
 ($0.5*(m-4-8.south east)+0.5*(m-5-8.north east)$);

\node[above=10pt of m-1-1] (top-1) {$M_1$};
\node[above=10pt of m-1-2] (top-2) {$M_2$};
\node[above=10pt of m-1-3] (top-3) {$M_3$};
\node[above=10pt of m-1-4] (top-4) {$M_4$};
\node[above=10pt of m-1-5] (top-5) {$M_5$};
\node[above=10pt of m-1-6] (top-6) {$M_6$};
\node[above=10pt of m-1-7] (top-7) {$M_7$};
\node[above=10pt of m-1-8] (top-8) {$M_8$};

\node[left=12pt of m-1-1] (left-1) {$M_1$};
\node[left=12pt of m-2-1] (left-2) {$M_2$};
\node[left=12pt of m-3-1] (left-3) {$M_3$};
\node[left=12pt of m-4-1] (left-4) {$M_4$};
\node[left=12pt of m-5-1] (left-5) {$M_5$};
\node[left=12pt of m-6-1] (left-6) {$M_6$};
\node[left=12pt of m-7-1] (left-7) {$M_7$};
\node[left=12pt of m-8-1] (left-8) {$M_8$};

\node[rectangle,above delimiter=\{] (del-top-1) at ($0.5*(top-1.south) +0.5*(top-4.south)$) {\tikz{\path (top-1.south west) rectangle (top-4.north east);}};
\node[above=10pt] at (del-top-1.north) {$Q-Snares$};
\node[rectangle,above delimiter=\{] (del-top-2) at ($0.5*(top-5.south) +0.5*(top-8.south)$) {\tikz{\path (top-4.south west) rectangle (top-6.north east);}};
\node[above=10pt] at (del-top-2.north) {$R-Snares$};

\node[rectangle,left delimiter=\{] (del-left-1) at ($0.5*(left-1.east) +0.5*(left-4.east)$) {\tikz{\path (left-1.north east) rectangle (left-4.south west);}};
\node[left=10pt] at (del-left-1.west) {$Q-Snares$};
\node[rectangle,left delimiter=\{] (del-left-2) at ($0.5*(left-5.east) +0.5*(left-8.east)$) {\tikz{\path (left-5.north east) rectangle (left-8.south west);}};
\node[left=10pt] at (del-left-2.west) {$R-Snares$};

\begin{pgfonlayer}{myback}

\foreach \element in {m-1-7,m-3-8,m-5-1,m-5-2,m-5-3,m-5-4,m-6-1,m-6-2,m-6-3,m-6-4,m-7-1,m-7-2,m-7-3,m-7-4,m-8-1,m-8-2,m-8-3,m-8-4}{
\highlight[pink]{\element}{\element}
}
\foreach \element in {m-2-7,m-3-6,m-4-5}{
\fhighlight[blue!10]{\element}{\element}
}
\end{pgfonlayer}

\end{tikzpicture}

\caption{Pairing matrix} \label{fig:M1}
\end{figure}