### 2017-07-13 paper/pm-1-matrix.tex

\begin{figure}
\centering
\begin{minipage}{0.45\linewidth}
  \hspace{-4ex}
\begin{tikzpicture}
\matrix [matrix of math nodes,left delimiter=(,right delimiter=),row sep=0.16cm,column sep=0.1cm] (m) {
\times & \times & \times &\times&\times&\times & \times & \times\\
\times & \times & \times&\times&\times&\times& 1 &\times \\
\times & \times & \times&\times&\times&1&\times &\times\\
\times & \times & \times&\times & 1 & \times &\times & \times\\
 \times & \times & \times & \times & \times & \times & \times & \times\\
 \times & \times & \times &\times & \times & \times & \times & \times\\
 \times & \times & \times & \times & \times & \times & \times & \times\\
 \times & \times & \times & \times & \times & \times&\times & \times\\};

\draw[dashed] ($0.5*(m-1-4.north east)+0.5*(m-1-5.north west)$) --
     ($0.5*(m-8-5.south east)+0.5*(m-8-4.south west)$);

\draw[dashed] ($0.5*(m-4-1.south west)+0.5*(m-5-1.north west)$) --
 ($0.5*(m-4-8.south east)+0.5*(m-5-8.north east)$);

\node[above=4pt of m-1-1] (top-1) {$M_1$};
\node[above=4pt of m-1-2] (top-2) {$M_2$};
\node[above=4pt of m-1-3] (top-3) {$M_3$};
\node[above=4pt of m-1-4] (top-4) {$M_4$};
\node[above=4pt of m-1-5] (top-5) {$M_5$};
\node[above=4pt of m-1-6] (top-6) {$M_6$};
\node[above=4pt of m-1-7] (top-7) {$M_7$};
\node[above=4pt of m-1-8] (top-8) {$M_8$};

\node[left=12pt of m-1-1] (left-1) {$M_1$};
\node[left=12pt of m-2-1] (left-2) {$M_2$};
\node[left=12pt of m-3-1] (left-3) {$M_3$};
\node[left=12pt of m-4-1] (left-4) {$M_4$};
\node[left=12pt of m-5-1] (left-5) {$M_5$};
\node[left=12pt of m-6-1] (left-6) {$M_6$};
\node[left=12pt of m-7-1] (left-7) {$M_7$};
\node[left=12pt of m-8-1] (left-8) {$M_8$};

\node[rectangle,above delimiter=\{] (del-top-1) at ($0.5*(top-1.south) +0.5*(top-4.south)$) {\tikz{\path (top-1.south west) rectangle (top-4.north east);}};
\node[above=10pt] at (del-top-1.north) {$Q-Snares$};
\node[rectangle,above delimiter=\{] (del-top-2) at ($0.5*(top-5.south) +0.5*(top-8.south)$) {\tikz{\path (top-4.south west) rectangle (top-6.north east);}};
\node[above=10pt] at (del-top-2.north) {$R-Snares$};

\node[rectangle,left delimiter=\{] (del-left-1) at ($0.5*(left-1.east) +0.5*(left-4.east)$) {\tikz{\path (left-1.north east) rectangle (left-4.south west);}};
\node[left=15pt,rotate=90,xshift=9mm] at (del-left-1.west) {$Q-Snares$};
\node[rectangle,left delimiter=\{] (del-left-2) at ($0.5*(left-5.east) +0.5*(left-8.east)$) {\tikz{\path (left-5.north east) rectangle (left-8.south west);}};
\node[left=15pt,rotate=90,xshift=9mm] at (del-left-2.west) {$R-Snares$};

\begin{pgfonlayer}{myback}

\foreach \element in {m-1-7,m-3-8,m-5-1,m-5-2,m-5-3,m-5-4,m-6-1,m-6-2,m-6-3,m-6-4,m-7-1,m-7-2,m-7-3,m-7-4,m-8-1,m-8-2,m-8-3,m-8-4}{
\highlight[pink]{\element}{\element}
}
\foreach \element in {m-2-7,m-3-6,m-4-5}{
\fhighlight[blue!10]{\element}{\element}
}
\end{pgfonlayer}

\end{tikzpicture}  
\end{minipage}
\begin{minipage}{0.45\linewidth}
  \tikzset{
    ln/.style  = { draw, thick, fill=black, circle, inner sep=0.3mm},
    sqloc/.style  = { draw, thick, inner sep=1.0mm },
    loc/.style    = { draw, circle, thick, inner sep=1.0mm },
  }
\begin{tikzpicture}
  \node[draw] (n0) {$n_1:\{0,1,2,3,4\}$};
  \node[draw,below left of=n0] (n1) {$n_1:\{0,1,2,3,4\}$};
  \node[draw] (n2) {$n_1:\{0,1,2,3,4\}$};

    \entity[yshift=-20mm] (E) {$n_1$ $[m_0 m_1 m_2 m_3 m_4]$};
    % \entity (E) {Node1\\[]};
    \binary[left=of E.north west, unique=2] (h) {};

       % \binary[left=of E.north west, unique=1-2] (w) {};

    \binary[left=of E.south west, unique=1-2](i) [below=6mm of h]{};
    \node[role name, at=(i.north),anchor=south] {[$m_{0} m_{1} m_{3}$]};
    \plays[] (E) to (i.east);
   % \plays(h) to (i);
%         label=below:\ormleft{uses] (u) {};
\node[role name,
at=(h.north),anchor=south] {[$m_{0} m_{1}$]};

    \entity[left=of h] (Rank) {$n_0:m_0 m_1 m_2 m_3 m_6$};

    \plays[mandatory] (E) to (h.east);
    \plays (h) to (Rank) (E);
    \plays[mandatory] (Rank) to (i.west);
    %to (u.east) (u) to (Car);

    \binary[right=of E.north east] (b) {};
    \node[role name,
at=(b.north),anchor=south] {[$m_{0} m_{1}$]};
       % \binary[right=of E.north east] (dd) {};

    \binary[right=of E.south east, unique,] (i) {};
    \node[role name,
at=(i.north),anchor=south] {[$m_{1} m_{2} m_{3}$]};
\binary[unique=1-2,skip unique=1-2] (xyz) [above=6mm of h] {};
% describes blue label above h Left uppermost
\node[role name,
at=(xyz.north),anchor=south] {[$m_{0} m_{1} m_{3}$]};

      \plays (Rank) to (xyz.west); 
    \plays[mandatory] (E) to (xyz.east);

    \entity[right=1.8 of E] (Date) {$n_2: m_1 m_2 m_3 m_6 m_8$};

    \plays[mandatory] (E) to (b.west); 
    \plays (E) to (i.west);
    \plays (b.east) to (Date);
    \plays[mandatory] (Date) to (i.east);

%    \entity[label={[constraint]\#=$n$}] {E};
%\draw (Rank) to node[vroles,label=has,label=below:\ormup{of}]{} (E);
%\node[constraint,anchor=north west,inner ysep=0]
%at (h.) {\ormbraces{a, b, c}};
%\draw (E) to node[roles,
%label=below:drives]{} (Rank);

\node[roles=2][,unique=1-2,skip unique=2] (abc) [below=11mm of E] {};
\node[role name,
at=(abc.north),anchor=south] {[$m_{2} m_{3}$]};
\plays[mandatory] (Rank.south) -- (abc.west);  
\plays (abc.east) -- (Date.south);

  %  \node[role name, anchor=south west] at (b.east) {[birthdate]};
  %  \node[role name, anchor=north west] at (i.east) {[hiredate]};

%     \rules at (-.4, -2) {
%          \node[rule=1] {{$m_{2}$} $m_{3}$};\\
%      };
\end{tikzpicture}
\end{minipage}

\caption{Pairing matrix} \label{fig:M1}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: