### 2017-07-13 paper/conclusion.tex
my changes

In the experiments, we have demonstrated the power of SMT solvers and the value
of careful encoding of the problems into SMT queries.
%
We manage to scale the tool upto the size that interests biologists.
%
However, there are many further search problems are of interest to
biologists about the VTSs.
%
For example, are there any k-connected graphs that can not be
converted into a VTS?
%
This encoding induces a quantifier alternation in the problem.
%
Therefore, a simple quantifier free encoding will not work.
%
We are planning to use QBF solvers to solve such queries or develop
iterative search algorithm for such queries.
%

% \ankit{interpret experiments (3-4 lines)}

% \srivas{Future work(quantifier problem etc)}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: