### 2017-07-18 paper/experiments.tex

\begin{table}[t]
\centering
% \def\arraystretch{1.6}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    {\multirow{2}{*}{\textbf{Version}}}  & 
    \multicolumn{2}{c|}{\textbf{Constraints}} & 
    \multicolumn{2}{c|}{\textbf{Graph connectivity}}  \\
   \cline{2-5}
   &  \textbf{Rest} & \textbf{Activity} & \textbf{\textbf{Least connected}} & \textbf{Sampling: Guarantee}\\ \hline

A. & \multirow{5}{*}{
     \makecell{ \ref{eq:f0}--\ref{eq:fuse2},\\
     \ref{eq:reach1},\ref{eq:reach2},\\
     \ref{eq:drop1}--\ref{eq:drop4}}
     } 
& \ref{eq:ann},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
B. & & \ref{eq:anb},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected & 3-connected \\  \cline{3-5}
D. & & \ref{eq:anb},\ref{eq:aeb} & 2-connected & 3-connected \\ \cline{3-5}
E. & & \ref{eq:ann},\ref{eq:aep} & No graph & No graph \\ \cline{3-5}
F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected & 4-connected \\ \hline
%G. & & \ref{eq:anb},\ref{eq:aeb},\ref{eq:ced} & 3-connected & 3-connected \\ \hline

% C_es :Self edges are allowed
% C_ed : Every edge is distinct 
  \end{tabular}
\caption{{\bf Activity regulation of molecules vs. graph connectivity.}}
\label{tab:grph}
\end{table}

% Given a system that follows the VTS, we are interested in its connectivity (graph) property.

% The basic idea is to constraint the variables such that vesicular
% transport rules are respected and using constraints (D0-D5) reason
% about the connectivity of the underlying structure (graph).

% We are interested in both least connectivity requirement (LRC) and least guarantee connectivity (LGC).
%LCR, that is required for a vesicular transport version to work and least connectivity that guarantees that vesicular transport system with that connectivity will always satisfies the rules.

% Whether a valid vesicular transport system is of certain connectiviity.   
%negation of the property .... STATE THE MEANING OF RESULT.

%Besides using the default Z3 solvers we have used for these experiments \ankit{Z3, picoSAT, Lingeling}. The performance of Z3 was ...
%
% \Ankit{In other solver variations ...}

We have implemented the encodings for each variants
using the python interface of $\zthree$ in a tool.
%
Our tool allows the user to choose a model and the size of the network
besides other parameters like connectivity and number of parallel
edges.
%
It uses Z3 Python interface to build the needed constraints and
applies $\zthree$ solver on the constraints to find a model (
a satisfying assignment that respects the constraints).
%
This tool also translates the satisfying model found by $\zthree$ into
a VTS and presents a visual output to the user in form of annotated graph.
%
% The output graph will satisfy the underlying rules of the system, so
% it always is a valid vesicle transport network.
%
% The resultant graph provides an overview of the underlying VTS
% where directed labeled edges and nodes provide the
% complete information about the system.
%
We also visually report the dropped edges to make the graph
disconnected, which gives information about the connectivity of the
graph.

% Dashed lines are used in the resultant visual network output to
% specify the dropped edges to make the graph disconnected, which gives
% information about the connectivity of the graph.
%\ashu{discuss table~\ref{tab:grph}}

In the table~\ref{tab:grph}, we also present the minimum connectedness needed for
the different variants after applying our tool for sizes from 2 to 10.
%
We haven't found any graph for the variant A with constraints Ann and Aen.
%
Replacing constraint on the node of Ann with Anb (version B) does not affect the outcome.
%
If we allow every present molecule to stay active (Ann) but constraint the edge
by a boolean function (Aeb) the resultant VTS has to be at least 3-connected.
%
If we constrain the activity of both the edge and the nodes with the boolean function (Anb and Aeb), 
the VTS has to be 2-connected.
%
Similarly, the results for the other cases are also presented in the table.

% Again there is no graph for the case where every molecule on the node is active, but the activity on the edge is driven by pairing inhibition (Aep).
% %
% If we replace the constraint on the node to be a boolean function (Anb) with Aep, we have found that 4-connected graphs satisfies all the constraints.

\input{tbl-qf-graph}

In Table~\ref{tab:qf-grabh}, we present the running times for the search of
VTSs of sizes 2 to 10 that satisfy the variants.
%
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation.
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
``!'' indicate that the constraints were unsatisfiable.
%
Hence with the use of this novel encoding, we are able to scale the
system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of ten compartments.
%
This additional scalability can provide us leverage to ask for related
open questions that were not previously possible, for example, what is
the minimum number of molecules required to satisfy the vesicular
traffic system in eukaryotes.
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.

% \input{cz-comparision}

% REQUIRE COMPARISION WITH CBMC RUN TIME ?

% %
% On this machine, the CBMC encoding ran out of memory (M/O) for N
% $\geq$ 6 whereas new novel encoding was able to handle the case fairly
% easily.
% %
% We think the main reasons for such an improvement are the following:
% \begin{enumerate}
% \item Direct encoding into the solver rather than using CFE (CBMC C front end), which treats front end variables as C variable and similarly other C overheads.
% \item Current novel encoding is fine-tuned to the structure of the problem.
% \item Improved encoding based on reachability rather than complete enumeration and non-determinism.
% \end{enumerate} 
% % 
% %

%table~\ref{tbl:qf-results}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: