### 2017-07-20 paper/experiments.tex
table 1 fixed

\begin{table}[t]
\centering
% \def\arraystretch{1.6}
  \begin{tabular}{|c|c|c|c|}
    \hline
    {\multirow{2}{*}{\textbf{Variant}}}  & 
    \multicolumn{2}{c|}{\textbf{Constraints}} & 
    \textbf{Graph connectivity}
    % \multicolumn{2}{c|}{\textbf{Graph connectivity}}
    \\
   \cline{2-4}
   &  \textbf{Rest} & \textbf{Activity} & \textbf{\textbf{Least connected}} % & \textbf{Sampling: Guarantee}
\\ \hline

A. & \multirow{5}{*}{
     \makecell{ \ref{eq:f0}--\ref{eq:fuse2},\\
     \ref{eq:reach1},\ref{eq:reach2},\\
     \ref{eq:drop1}--\ref{eq:drop4}}
     } 
     & \ref{eq:ann},\ref{eq:aen} & No graph     \\\cline{3-4}
B. & & \ref{eq:anb},\ref{eq:aen} & No graph     \\\cline{3-4}
C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected  \\\cline{3-4}
D. & & \ref{eq:anb},\ref{eq:aeb} & 2-connected  \\\cline{3-4}
E. & & \ref{eq:ann},\ref{eq:aep} & No graph     \\\cline{3-4}
F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected  \\\hline

% C_es :Self edges are allowed
% C_ed : Every edge is distinct 
  \end{tabular}
\caption{{\bf Activity regulation of molecules vs. graph connectivity.}}
\label{tab:grph}
\end{table}

% Given a system that follows the VTS, we are interested in its connectivity (graph) property.

% The basic idea is to constraint the variables such that vesicular
% transport rules are respected and using constraints (D0-D5) reason
% about the connectivity of the underlying structure (graph).

% We are interested in both least connectivity requirement (LRC) and least guarantee connectivity (LGC).
%LCR, that is required for a vesicular transport version to work and least connectivity that guarantees that vesicular transport system with that connectivity will always satisfies the rules.

% Whether a valid vesicular transport system is of certain connectiviity.   
%negation of the property .... STATE THE MEANING OF RESULT.

%Besides using the default Z3 solvers we have used for these experiments \ankit{Z3, picoSAT, Lingeling}. The performance of Z3 was ...
%
% \Ankit{In other solver variations ...}

We have implemented the encodings for each variants
using the python interface of $\zthree$ in a tool.
%
Our tool allows the user to choose a model and the size of the network
besides other parameters like connectivity and number of parallel
edges.
%
It uses Z3 Python interface to build the needed constraints and
applies $\zthree$ solver on the constraints to find a model (
a satisfying assignment that respects the constraints).
%
This tool also translates the satisfying model found by $\zthree$ into
a VTS and presents a visual output to the user in form of annotated graph.
%
% The output graph will satisfy the underlying rules of the system, so
% it always is a valid vesicle transport network.
%
% The resultant graph provides an overview of the underlying VTS
% where directed labeled edges and nodes provide the
% complete information about the system.
%
We also visually report the dropped edges to make the graph
disconnected, which gives information about the connectivity of the
graph.

% Dashed lines are used in the resultant visual network output to
% specify the dropped edges to make the graph disconnected, which gives
% information about the connectivity of the graph.
%\ashu{discuss table~\ref{tab:grph}}

In the table~\ref{tab:grph}, we also present the minimum connectedness needed for
the different variants after applying our tool for sizes from 2 to 10.
%
We haven't found any graph for the variant A with constraints Ann and Aen.
%
Replacing constraint on the node of Ann with Anb (version B) does not affect the outcome.
%
If we allow every present molecule to stay active (Ann) but constraint the edge
by a boolean function (Aeb) the resultant VTS has to be at least 3-connected.
%
If we constrain the activity of both the edge and the nodes with the boolean function (Anb and Aeb), 
the VTS has to be 2-connected.
%
Similarly, the results for the other cases are also presented in the table.

% Again there is no graph for the case where every molecule on the node is active, but the activity on the edge is driven by pairing inhibition (Aep).
% %
% If we replace the constraint on the node to be a boolean function (Anb) with Aep, we have found that 4-connected graphs satisfies all the constraints.

\input{tbl-qf-graph}

In Table~\ref{tab:qf-grabh}, we present the running times for the search of
VTSs of sizes 2 to 10 that satisfy the variants.
%
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation.
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
``!'' indicate that the constraints were unsatisfiable.
%
Hence with the use of this novel encoding, we are able to scale the
system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of ten compartments.
%
% \ankit{
% This additional scalability can provide us leverage to ask for related
% open questions that were not previously possible, for example, what is
% the minimum number of molecules required to satisfy the vesicular
% traffic system in eukaryotes.}
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.

% \input{cz-comparision}

% REQUIRE COMPARISION WITH CBMC RUN TIME ?

% %
% On this machine, the CBMC encoding ran out of memory (M/O) for N
% $\geq$ 6 whereas new novel encoding was able to handle the case fairly
% easily.
% %
% We think the main reasons for such an improvement are the following:
% \begin{enumerate}
% \item Direct encoding into the solver rather than using CFE (CBMC C front end), which treats front end variables as C variable and similarly other C overheads.
% \item Current novel encoding is fine-tuned to the structure of the problem.
% \item Improved encoding based on reachability rather than complete enumeration and non-determinism.
% \end{enumerate} 
% % 
% %

%table~\ref{tbl:qf-results}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: