### 2018-04-15 papers/synth/abstract.tex

Vesicle Traffic Systems are the transport mechanisms inside the cells.
%
Different compartments are viewed as nodes and transport channels are
nodes between the graphs.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: