### 2018-04-15 papers/synth/abstract.tex
abstract in

Vesicle Traffic Systems(VTS) are the material transport mechanisms
among the compartments inside the cells.
%
The compartments are viewed nodes that labeled with the containing
material and the transport channels are labeled edges between the
graphs.
%
Understanding VTS is an ongoing area of research.
%
There may be undiscovered edges or nodes.
%
The labels may not have been fully observed.
%
Furthermore, it is matter of debate what properties
these systems should have, such as stability, i.e., every
chemical that is leaving a compartment comes back. 
%
In these paper, we present the encoding of the above questions
into the QBF satisfiability problems.
%
We have implemented the encodings in a highly configurable tool
and  apply to a couple of found-in-nature VTSs and 
several synthetic graphs.
%
Our results demonstrates that our method can scale upto the
graphs of interest.
%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: