### 2018-04-22 papers/synth/conclusion.tex
intro

The study of dynamic behaviour of VTSs may be the focus of future work.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: