### 2018-04-30 papers/synth/conclusion.tex
encoding finished

In this paper, we presented the encoding of the synthesis problems
that may rise from VTSs.
%
We demonstrated that our tool scale up to the relevant sizes of the
VTSs. 
%
We will take this tool to the biologists and develop wet experiments 
that may validate some synthesis results from the tool.
%
Our model of VTSs are static graphs.
%
In future, we will study the dynamic behaviours of VTSs.
%
It will allow us to predict behaviors after the perturbations in the VTSs
and more ways to test the predicted synthesis results.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: