### 2018-05-03 papers/synth/related.tex

In recent years, there has been a wide range of methods
developed for the similar synthesis problems~\cite{sktech,ceSynth,exampleSynth}.

There are a few examples of using synthesis technique in biology especially in gene regulatory networks~\cite{shavit2016automated, fisher2015synthesising}. A very recent work~\cite{fisher2015synthesising} synthesize executable gene regulatory networks from single-cell gene expression data. 
Synthesis technique is also used in optimal synthesis for chemical reaction networks~\cite{cardelli2017syntax}. the~\cite{fisher2015synthesising} uses constraint (satisfiability) solving techniques for the synthesis whereas ~\cite{shavit2016automated} uses SMT for synthesis. The queries contain quantifiers so we have we employ QBF solving for the solving the synthesis problem. The paper~\cite{cardelli2017syntax} uses a template-guided approach which is very close to our work. To our best knowledge this is the first application of synthesis in VTS.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: