### 2018-05-04 papers/synth/fig-yeast-vts.tex
asd

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[->,>=stealth',auto,node distance=5cm,
  thick,main node/.style={rectangle,draw,font=\sffamily\Large\bfseries}]
  \node[main node,,text width=2.5cm] (golgi) [] {GOLGI {\small \^{Qa2} \^{Qb2} \^{Qc2} \^{Qa5} \^{Qb5} \^{Qc5}}};
  \node[main node,text width=2.5cm] (pvac) [right of=golgi] {PRE-VAC {\small \^{Qa3} \^{Qb2} \^{R3} \^{R4}}};
  \node[main node,text width=2cm] (vac) [right of=pvac] {VAC {\small \^{Qa4} \^{Qb2} \^{R3} \^{R4}}};
  \node[main node,text width=2.5cm] (pm_up) [above right of=golgi] {PM {\small \^{Qa1} \^{Qbc1}}};
  \node[main node] (pm_down) [below right of=golgi] {PM {\small \^{Qa6} \^{Qb6} \^{R6} \^{R4}}};

  \path (golgi) edge[bend right] node [right] {\^{R1}} (pm_up);
  \path (pm_up) edge[bend right] node [right] {\^{R1}} (golgi);

  \path (golgi) edge[bend right] node [right] {\^{Qc6}} (pm_down);
  \path (pm_down) edge[bend right] node [right] {\^{R1}} (golgi);

  \path (golgi) edge[] node [below] {\^{Qc3}} (pvac);
  \path (pvac) edge[bend left=20] node [above] {\^{Qc4}} (vac);

  \end{tikzpicture}
  \caption{Yeast VTS}
  \label{fig:mukund-vts}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: