### 2018-05-05 papers/synth/example.tex
sad

\input{fig-mukund-vts}

In figure~\ref{fig:mukund-vts}, we present the VTS for mammalian cells
obtained by studying the literature.
%
The

% and hollow arrows represent maturation edges.

We have used our tool to synthesize possible fixes to complete the
graph. Based on the variation we were able to synthesis the this graph
within 600s. We have created a rank of minimum edits to complete the
graph. Our experiments suggest that the synthesis problems are
solvable by modern solvers and the synthesis technology may be useful
for biological research.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: