### 2018-05-05 papers/synth/example.tex

\input{fig-mukund-vts}

In figure~\ref{fig:mukund-vts}, we present the VTS for mammalian cells
obtained by studying the literature.
%
The VTS has six nodes and 20+ molecules.
%
We can easily check that stability condition is not satisfied for
many molecules for example ???.
%
It is clear that there are several unknowns about the VTS.
%
Please read appendix~\ref{sec:ex-vts} for more biological details.

% and hollow arrows represent maturation edges.

We have used our tool to synthesize possible fixes to complete the
graph.
%
Based on the variation we were able to synthesis the this graph
within ??s.
%
We added ?? edges to make the graph stable.
%
We have created a rank of minimum edits to complete the
graph.
%
Our experiments suggest that the synthesis problems are
solvable by modern solvers and the synthesis technology may be useful
for biological research.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: