### 2018-05-05 papers/synth/example.tex

% \input{fig-mukund-vts}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: