### 2018-06-14 papers/journal/abstract.tex

Vesicle Traffic Systems(VTSs) are the material transport 
mechanisms among the compartments inside the biological cells.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: