### 2018-11-09 papers/journal/abstract.tex
Updated the directory from the Overleaf directory, added encodings from CBMC paper/ PLOS paper

Vesicle Traffic Systems (VTSs) are the material transport
mechanisms among the compartments inside the biological cells.
%
The compartments are viewed as nodes that are labeled with the
containing chemicals and the transport channels
are similarly viewed as labeled edges between the nodes.
%
Several interesting questions about VTS translates to combinatorial
search and synthesis problem. 
%
We present novel SAT, SMT and QBF based encodings of the properties
over vesicle traffic systems and a tool that searches for the networks
that satisfies the properties using these solvers. 
%
Understanding VTSs is an ongoing area of research and for many cells
they are partially known...
%
In our experiments, we show that our tool can search for network of sizes that are considered
to be relevant by biologists.

% Vesicle Traffic Systems(VTSs) are the material transport 
% mechanisms among the compartments inside the biological cells.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: