### 2018-11-15 papers/journal/qbf-table.tex

\begin{table}[t]
	\centering
	    \def\arraystretch{1.1}
%	\caption{{\bf SAT and QBF implementation comparison.}}
%	\adjustbox{max height=\dimexpr\textheight-5.5cm\relax,
%		max width=\textwidth}{
		\begin{tabular}[t]{|c|c|c|c|c|c|}\hline

			{\multirow{2}{*} {Sr.No}}  & \multicolumn{2}{c|}{Regulation} & \multicolumn{3}{c|}{Universal condition}

		%	\\\cline{2-7}

%			{} & \multicolumn{2}{c|}{\textbf{2-connected} } & \multicolumn{2}{c|}{\textbf{3-connected} } & \multicolumn{2}{c|}{\textbf{2-connected}}  &  \multicolumn{2}{c|}{\textbf{4-connected}}

			%{} & \multicolumn{2}{c|}{2-connected} & \multicolumn{2}{c|}{3-connected} & \multicolumn{2}{c|}{2-connected}  &  \multicolumn{2}{c|}{4-connected}

			\\\cline{2-6}
			{} & {Node} & {Edge} & {Connectivity} & {SAT-time} & {QBF-time}

			\\\hline
			1. & Bf & Bf & 3C & 2.12 & 1.89  \\\hline
			2. & None & Bf  & 3C & 7.12 & 4.89  \\\hline
			3. & Bf & SNARE inhib & 4C  & M/O & 7.66 \\\hline
			4. & None & SNARE inhib  & No graph & M/O & 48.35 \\\hline
			5. & Bf & None  & No graphs & M/O & 890.84 \\\hline
			6. &  None & None  & No graphs & M/O & M/O \\\hline

	\end{tabular}
%}
	\caption{ SAT and QBF implementation comparison (in secs.)}
	\label{tab:satqbf-graph}
\end{table}