### 2019-01-13 papers/journal/qbf-table.tex
Universal condition table restructure

%\begin{table}[t]
%	\centering
%	    \def\arraystretch{1.2}
%%	\caption{{\bf SAT and QBF implementation comparison.}}
%%	\adjustbox{max height=\dimexpr\textheight-5.5cm\relax,
%%		max width=\textwidth}{
%		\begin{tabular}[t]{ c|c c|c }\hline
%			
%			{\multirow{2}{*} {Sr.No}}  & \multicolumn{2}{c|}{Regulation} &
%			{\multirow{2}{*} {Universal condition}} 
%			
%		%	\\\cline{2-7}
%			
%%			{} & \multicolumn{2}{c|}{\textbf{2-connected} } & \multicolumn{2}{c|}{\textbf{3-connected} } & \multicolumn{2}{c|}{\textbf{2-connected}}  &  \multicolumn{2}{c|}{\textbf{4-connected}}
%
%			%{} & \multicolumn{2}{c|}{2-connected} & \multicolumn{2}{c|}{3-connected} & \multicolumn{2}{c|}{2-connected}  &  \multicolumn{2}{c|}{4-connected}
%			
%			\\
%			\cline{2-3}
%			{} & {Node} & {Edge} 
%			% &  {QBF-Scalable}
%			
%			\\\hline
%			1. & Bf & Bf & 3-Connected   \\\hline
%			2. & None & Bf  & 3-Connected   \\\hline
%			3. & Bf & SNARE inhib & 4-Connected  \\\hline
%			4. & None & SNARE inhib   & No graphs  \\\hline
%			5. & Bf & None  & No graphs   \\\hline
%			6. &  None & None  & No graphs   \\\hline
%			
%	\end{tabular}
%%}
%	\caption{ QBF result and timings (in secs.)}
%	\label{tab:qbf-graph}
%\end{table}
%
%

\begin{table}[t]
	\centering
	\def\arraystretch{1.2}
	\begin{tabular}{ c|c|c|c }
		\hline
		{\multirow{2}{*}{Variant}}  &
		\multicolumn{2}{c|} {Constraints} & 
		%	{\multirow{2}{*}{\textbf{Results}}}
		{\multirow{2}{*}{Graph connectivity }} 
		\\
		\cline{2-3}
		&  Rest & Activity & 
		\\ \hline

		A. & \multirow{5}{*}{
			\makecell{ \texttt{Consistancy}, \\  \texttt{Stability}, \\
				 \texttt{Connected(k)}  }
		}
		& \ref{eq:ann},\ref{eq:aen} & No graph   \\
		B. & & \ref{eq:anb},\ref{eq:aen} & No graph \\
		C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected \\
		D. & & \ref{eq:anb},\ref{eq:aeb} & 3-connected \\
		E. & & \ref{eq:ann},\ref{eq:aep} & No graph    \\
		F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected \\\hline

		% C_es :Self edges are allowed
		% C_ed : Every edge is distinct 
	\end{tabular}
	\caption{ QBF result for graph connectivity}
	\label{tab:qbf-graph}
\end{table}