### 2019-01-13 papers/journal/qbf-table.tex
Fancy table everywhere

\begin{table}[t]
	\centering
	\def\arraystretch{1.2}
%	\begin{tabular}{ c|c|c|c }
%		\hline
%		{\multirow{2}{*}{Variant}}  &
%		\multicolumn{2}{c|} {Constraints} & 
%		%	{\multirow{2}{*}{\textbf{Results}}}
%		{\multirow{2}{*}{Graph connectivity }} 
%		\\
%		\cline{2-3}
%		&  Rest & Activity & 
%		\\ \hline
%		
%		A. & \multirow{5}{*}{
%			\makecell{ \texttt{Consistancy}, \\  \texttt{Stability}, \\
%				 \texttt{Connected(k)}  }
%		}
%		& \ref{eq:ann},\ref{eq:aen} & No graph   \\
%		B. & & \ref{eq:anb},\ref{eq:aen} & No graph \\
%		C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected \\
%		D. & & \ref{eq:anb},\ref{eq:aeb} & 3-connected \\
%		E. & & \ref{eq:ann},\ref{eq:aep} & No graph    \\
%		F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected \\\hline
%		
%		% C_es :Self edges are allowed
%		% C_ed : Every edge is distinct 
%	\end{tabular}
	\begin{tabular}{ c|c c c|c }
	\hline
	{\multirow{3}{*}{Variant}}  &
	\multicolumn{3}{c|} {Constraints} & 
	%	{\multirow{2}{*}{\textbf{Results}}}
	{\multirow{3}{*}{Graph connectivity} } 
	\\
	\cline{2-4}
	& 	{\multirow{2}{*}{Rest} }  & 	\multicolumn{2}{c|} {Activity} & 
	\\ 
	\cline{3-4}
	&  & Node  & Edge & 
	\\\hline

	A. & \multirow{5}{*}{
					\makecell{ \texttt{Consistancy}, \\  \texttt{Stability}, \\
						 \texttt{Connected(k)}  }
	}
	& \ref{eq:ann} & \ref{eq:aen} & No graph  \\
	B. & & \ref{eq:anb} & \ref{eq:aen} & No graph \\
	C. & & \ref{eq:ann} &  \ref{eq:aeb} &  3-connected \\
	D. & & \ref{eq:anb} & \ref{eq:aeb} & 3-connected \\
	E. & & \ref{eq:ann} & \ref{eq:aep} & No graph  \\
	F. & & \ref{eq:anb} &  \ref{eq:aep} & 4-connected \\\hline

	% C_es :Self edges are allowed
	% C_ed : Every edge is distinct 
\end{tabular}
	\caption{ QBF result for graph connectivity}
	\label{tab:qbf-graph}
\end{table}