### 2019-04-03 papers/journal/synth-table.tex

\begin{sidewaysfigure}[t]
	\centering
	\begin{tabular}[t]{|c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}C{4cm}@{}|}\hline
		% \begin{table}[t]
		%   \centering
		%   \begin{tabular}[t]{|c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|}\hline
		{\multirow{2}{*} \textbf{}}  & \multicolumn{2}{c|}{\textbf{Add}} & \multicolumn{2}{c|}{\textbf{Add}} & \multicolumn{2}{c|}{\textbf{Learning NNF}}  &  \multicolumn{2}{c|}{\textbf{Learning}} &  \multicolumn{2}{c|}{\textbf{Add/Delete}} \\
		{\multirow{2}{*} \textbf{Table a}}  & \multicolumn{2}{c|}{\textbf{edge}} & \multicolumn{2}{c|}{\textbf{molecules}} & \multicolumn{2}{c|}{\textbf{(only $\land$ and $\lor$)}}  &  \multicolumn{2}{c|}{\textbf{k-CNF}} &  \multicolumn{2}{c|}{\textbf{parts}} \\
		\cline{2-11}
		{} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} \\
		\hline

		plos1-dia[3C]& 0.02 &$\infty$& 0.37 &$\infty$& 0.46 & $\infty$ & 0.63 &$\infty$& 0.21 & -1 E, -1AE, -1AN. +1E, +1N. \\\hline
		plos2-dia[4C] & 2.02 & 0   & 2.38 & 0  & 4.27  & 0 & 5.94 & 0 &  3.30 & 0 \\\hline
		sub-mammal[3C]  & 4.61 & 1 E  & 5.98 & 7 PE & 27.33 & 1 E & 14.75 & 1 E & 6.82  & -1E, -2PE, -4AN. +1 E, +5PE, +4N, +3AN, +3AE. \\\hline
		node4[3C]  & 5.93  & 1 E   &  13.72 & 12 PE  &  5.09  & $\infty$ & 4.30 &$\infty$& 7.21  & -2 E, -2PE, -1 N, -1AN, -1 AE. +18N, +8E, +1PE.\\\hline
		%   yeast-graph[3C]   & 95.016    & 1 E   & 94.520   & 1 E   & 169.430  & 1 E & 172.35   & 1E   & 107.43  &  -1 E, -2 N, -2 AE, -2 AN. +2 E, 12 PE, 7 N. \\\hline
		yeast-graph[3C]   & 105.15    & 2 E  &   247.04  & 15 PE   & 1381.24  & 2 E  & 541.66   & 2 E & 326.09  &  -1E, -1N, -1AE, -1AN, -1PE. +2E, +11PE, +20N. \\\hline
	   	node6[3C]  &  454.98  & 2 E  &  1209.30    & 21 PE    &  T/O  & N/A      &  T/O   &  N/A   &  T/O     & N/A \\\hline
		mammal-graph[3C]  &  T/O     & N/A  &  T/O     & N/A    &  T/O         & N/A      &  T/O    &  N/A    &  T/O     & N/A\\\hline
		%    & 0.0    & 0.0    & 0.0    & 0.0    & 0.0         & 0.0      & 0.0   & 0.0    & 0.0    & 0.0\\\hline
	\end{tabular}
	% \caption{Run-times for searching for models (in secs). \#C  stands for minimum changes.
	% Time is reported in seconds.}
	% \label{tab:qf-graph}
	% \end{table}
	\begin{tabular}[t]{|c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}c@{}|@{}C{4cm}@{}|}\hline
		{\multirow{2}{*} \textbf{}}  & \multicolumn{2}{c|}{\textbf{Add}} & \multicolumn{2}{c|}{\textbf{Add}} & \multicolumn{2}{c|}{\textbf{Learning NNF}}  &  \multicolumn{2}{c|}{\textbf{Learning}} &  \multicolumn{2}{c|}{\textbf{Add/Delete}} \\
		{\multirow{2}{*} \textbf{Table b}}  & \multicolumn{2}{c|}{\textbf{edge}} & \multicolumn{2}{c|}{\textbf{molecules}} & \multicolumn{2}{c|}{\textbf{(only $\land$ and $\lor$)}}  &  \multicolumn{2}{c|}{\textbf{k-CNF}} &  \multicolumn{2}{c|}{\textbf{parts}} \\
		\cline{2-11}
		{} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} & {\textbf{Time}} & {\textbf{\#C}} \\
		\hline

		plos1-dia & 0.04&$\infty$& 0.31 &$\infty$& 0.20 & $\infty$ & 0.27&$\infty$& 3.47 & -1E, -1PE, - 1N, -1PE. +1AE, +1PE, +1N\\\hline
		plos2-dia & 3.97 & 0 &  2.26 & 0  & 5.46 & 0 & 5.35 & 0 & 3.24 & 0 \\\hline
		sub-mammal & 3.22 & 1 E  & 4.18 & 7 PE  & 24.73 & 1 E  & 9.21 & 1 E & 5.36  & -1E, -2PE, -4AN. +1 E, +5PE, +4N, +3AN, +3AE \\\hline
		node4  & 4.01  & 1 E  & 10.19  & 12 PE & 4.26  & $\infty$ & 4.13 &$\infty$&  6.89  & -2 E, -2PE, -1N, -1AN, -1AE. +18N, +8E, +1PE \\\hline
		yeast-graph & 38.68  & 2 E  &   127.45  & 15 PE   & 1002.41  & 2 E  & 434.08   & 2 E & 197.82  &  -1E, -1N, -1AE, -1AN, -1PE. +2E, +11PE, +20N. \\\hline
		node6  &  235.97  & 2 E  &  549.25    & 21 PE    &  T/O  & N/A      &  T/O   &  N/A   &  T/O     & N/A \\\hline
		mammal-graph   &  T/O     & N/A  &  T/O    & N/A    &  T/O  & N/A      &  T/O   &  N/A   &  T/O     & N/A\\\hline
	\end{tabular}
	\caption{Run-times for synthesis queries. \#C stands for minimum changes in the synthesized VTS in comparison with the given partial VTS. Time is reported in seconds. (a) The solver used is DepQBF (b) The solver used is Z3. The sub-mammal is a subgraph of the complete mammal-graph. In the Add/Delete parts column, ‘+’n sign is used to show the addition of n number of the molecules, similarly ‘-’n is used to show the removal of n number of molecules. In the table, N is node labels, AN is active node molecules, E is edges, PE is molecule presence on the edge and AE is active molecules on the edge. The [kC] stands for k graph connectedness which is part of only DepQBF experiments.}

	% \caption{Run-times for searching for Z3 models (in secs). \#C  stands for minimum changes.
	% Time is reported in seconds.}
	\label{tab:synth-graph}
\end{sidewaysfigure}