### 2019-04-11 papers/journal/conclusion.tex
Conclusion section

\noindent 
We have established a connection between vesicle transport graph connectedness and underlying rules of SNARE pairing and regulation. 
%
The constraints implied by SNARE protein properties translates into
mathematically elegant properties of the transport graph, namely its connectivity.
%
Our approach demonstrates the power of SMT solvers in a {\em novel} application and the value of careful encoding of the problems into SMT queries. 
We manage to scale the tool up to the size that makes the tool biologically relevant.
%

However, the problem of searching for all $k$-connected VTSs induces quantifier alternation in the encodings.
%
Therefore, a simple call to SAT solver may not work efficiently. 
%
We used QBF (quantified Boolean
formulas) solvers for such queries.
%
One may also be interested in counting the number of graphs that satisfy
a given property. 
%
Exact counting of the graphs using SAT solvers may prove to be very
difficult. 
%
We are planning to employ approximate counting methods for finding the solution.
%
Our results contain many predictions which can in principle be tested through cell-biological experiments.

We also presented encodings of the synthesis problems that may arise from VTSs.
%
We demonstrated that our tool based on the encodings
scale up to the relevant sizes of the VTSs for some synthesis queries.
%
Our tool timed out on larger examples.
%
We are working to improve the performance of our tool.
%
%
With the tool, we made an attempt to present multiple new testable predictions. 
%
We will take this tool to the biologists and develop wet experiments that may
validate some synthesis results from the tool.
%
These predictions may very likely increase to hasten discoveries in this field
alongside pushing forward the use of formal methods in the biological field.
%

Since our current knowledge of VTSs is not extensive and there are many missing parts,
there is a huge space of potential predictions.
%
We also developing methods to prioritize predictions considering addition biological
information and using a refined model.
%
For example, in this work, we have only analyzed the VTS as static graphs. A major aspect of the biological cell’s VTS that the current model is missing is \textit{compartment maturation}, imbalances in incoming and outgoing molecular flux on vesicles.
%
In the future, we will study the dynamic behaviors of VTSs.
%
%Maturation is evident in the Golgi in the endocytic system. 
%
%Including this behavior of the compartments in the model would allow it to make more accurate predictions.

%Our model of VTSs is static graphs.
%
% There are a few aspects in the modeling that we have relaxed.
%  \item What  needs  to  be  done  to  use  these  tools  to  real  world  BIO problems. In what aspect you want more, where it lagged.
%\end{enumerate}
%\textbf{Encode Dynamic behaviour of the VTS}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: