### 2019-04-11 papers/journal/conclusion.tex
conclusion

\noindent 
To conclude, we have established a connection between vesicle transport graph connectedness and underlying rules of SNARE pairing and regulation. 
%
The constraints implied by SNARE protein properties translates into
mathematically elegant properties of the transport graph, namely its connectivity.
%
Our approach demonstrates the power of SMT solvers in a {\em novel} application and the value of careful encoding of the problems into SMT queries. 
We manage to scale the tool up to the size that makes the tool biologically relevant.
%

The problem of searching for all $k$-connected VTSs induces quantifier alternation in the encodings.
%
Therefore, a simple call to SAT solver may not work efficiently. 
%
We used QBF (quantified Boolean
formulas) solvers for such queries.
%
Our connectivity results contain many predictions which can in principle be tested through cell-biological experiments. 
%Our connectivity results for VTS contains many predictions which can in
%principle be tested through cell-biological experiments. 
%%
%However, our current knowledge of the actual vesicle transport graph of a cell is very poor: it is likely that there are many vesicle transport routes that are still uncharacterized. Even in those instances where vesicles are known to exist, the precise molecular drivers of their creation and fusion are often unknown.
%%
%Therefore, a direct verification of our predictions based on graph connectedness alone is not likely. 
%%
%However, more indirect predictions based on how SNAREs are regulated can be
%explored. 
%%
%In this context, the Sec/MUNC proteins or SNARE longin domains are interesting
%targets of study~\cite{van2010one,rossi2004longins}
One may also be interested in counting the number of graphs that satisfy
a given property. 
%
Exact counting of the graphs using SAT solvers may prove to be very
difficult. 
%
We are planning to employ approximate counting methods for finding the solution.
%

We also presented encodings of the synthesis problems that may arise from VTSs.
%
We demonstrated that our tool based on the encodings
scale up to the relevant sizes of the VTSs for some synthesis queries.
%
Our tool timed out on larger examples.
%
We are working to improve the performance of our tool.
%
%
With the tool, we made an attempt to present multiple new testable predictions. 
%
We will take this tool to the biologists and develop wet experiments that may
validate some synthesis results from the tool.
%
These predictions may very likely increase to hasten discoveries in this field
alongside pushing forward the use of formal methods in the biological field.
%

Since our current knowledge of VTSs is not extensive and there are many missing parts,
there is a huge space of potential predictions.
%
We also developing methods to prioritize predictions considering addition biological
information and using a refined model.
%
For example, in this work, we have only analyzed the VTS as static graphs. A major aspect of the biological cell’s VTS that the current model is missing is \textit{compartment maturation}, imbalances in incoming and outgoing molecular flux on vesicles.
%
In the future, we will study the dynamic behaviors of VTSs.
%
%Maturation is evident in the Golgi in the endocytic system. 
%
%Including this behavior of the compartments in the model would allow it to make more accurate predictions.

%Our model of VTSs is static graphs.
%
% There are a few aspects in the modeling that we have relaxed.
%  \item What  needs  to  be  done  to  use  these  tools  to  real  world  BIO problems. In what aspect you want more, where it lagged.
%\end{enumerate}
%\textbf{Encode Dynamic behaviour of the VTS}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: