### 2019-09-09 papers/journal/2019_ACTA_Rebuttal_Rd2.txt
renamed it to .txt

% Reviewer 1:
    No Comments!

% Reviewer 3:

> I believe that most issues highlighted previously have been 
  addressed adequately. In particular, the structure of the 
  revised manuscript has been improved significantly. 
  I still found the paper challenging to read, but this is 
  largely due to the complexity of the biological system and 
  the necessary encoding details presented in the manuscript.

  Pt #1:
-------- Handle by Mukund and Ashutosh --------------
  One point that I think requires further attention is the 
  motivation for using connectedness and the explanation of 
  this property, which is at the core of the proposed approach. 
  I did not find the additional explanations provided in the 
  response and the revised manuscript sufficient. Simply stating 
  that "The graph k-connectedness furnishes precisely" a test to 
  determine if information is missing from the model and that 
  "If the data […] does not have a certain degree of connectivity, 
  this implies that some biological data has been missed" does not 
  justify connectedness as a fundamental and required biological 
  property that all valid models should respect. The intuition of 
  connectedness as a robustness property provided elsewhere in the 
  manuscript helps but it is still not clear what evidence supports 
  the fact that VTSs are indeed robust. In addition, the problem as 
  formulated does not seem to seek the most robust networks as 
  explained in the following.

  Pt #2: 
-------- Handle by Ankit --------------
  I also found the formulation of the main problems with respect 
  to connectedness somewhat confusing. For example, in Sec. 5.1.1 
  and Sec. 5.1.2 the problems are formulated as a search for VTS G 
  such that G is not k-connected. Yet, Sec. 5.1 specifies that 
  "In case the formula is satisfiable the procedure terminates and 
   we report the value of k", which seems contradictory. Finally, 
   it is not clear why the goal is to find the minimum graph 
   connectivity, given that larger connectivity corresponds to a 
   more robust network. I was expecting an approach involving 
   incrementing k to find the smallest value that makes the formula 
   unsatisfiable, thus identifying the maximum graph connectedness 
   possible for a given VTS with all other constraints being 
   satisfied. However, to the best of my understanding, this is not 
   what the paper is currently proposing, or else further  
   clarifications are needed.

  Pt #3: 
-------- Handle by Ashutosh --------------   
  Another point that should be revised further is around 
  clarifying the contributions of this paper. The response 
  document summarizes nicely that, together with additional 
  QBF encoding, "this paper brings together all the results 
  with extended experiments and more detailed expositions for 
  the benefit of both biological and computation community." 
  However, I did not find a similar concise statement of the 
  contributions in the text. In addition, some of the stated 
  advances over previous approaches seem incremental and minor. 
  For example, for the previous CBMC-based method it is stated 
  that "...the scalability of the method was insufficient. 
  The tool was able to handle the system with just eight 
  compartments". Yet, we learn that "A typical eukaryotic 
  cell contains 9 organelles" and even one of the methods 
  presented in this paper fails for some of the experimental 
  evaluations with 7 or more compartments. The contributions 
  of this work, perhaps in the integration and evaluation of
  several methods reported previously, should be stated clearly 
  and accurately in the main text.

  Pt #4: 
-------- Handle by Ankit and Ashutosh--------------  
  A final minor point is about the use of "Synthesis" as 
  an encoding in Fig 4 and the relevant explanations. 
  Since the other problems list verification techniques 
  and solvers (e.g. BMC, SMT, QBF) and no novel synthesis 
  algorithms were developed for this problem, it would be 
  more appropriate to specify QBF as the encoding used for 
  the "VTS synthesis" problem.

% Reviewer 4:

  Pt #1:
-------- Handle by Somya--------------  
  Point 6 on page 3 is the same as point 4.

  Pt #2:
-------- Handle by Somya--------------  
  I would argue still that the biological section has 
  unnecessary detail and could be simplified further.

  Pt #3:
-------- Handle by Somya--------------  
  That the steady-state condition corresponds to the 
  biological meaning of homeostasis is repeated. This 
  only needs to be mentioned once.

  Pt #4:
-------- Handle by Ankit--------------  
   There is overlap in the 3rd and 6th paragraphs of 
   the 'Related Work' section, which refer to the use 
   of SAT/SMT in GRNs.

Minor points: 
-------- Handle by Lakhsman--------------   
  Small typos / grammatical errors. There are a number 
  of these throughout the text, some of which are listed 
  below (this is not exhaustive, so I recommend carefully 
  reading the manuscript to ensure clarity):

  Page 2, lines 11-12: Understanding how membrane traffic 
  functions *is*, therefore,…

  Page 2, line 13: Cell biologists have made… (not 'The')

   Page 3, line 14: "…parts of the traffic system that 
   *are* unchanging…"

   Page 4, line 6: "We also worked on the synthesis of the 
   unknown pieces *that* may be assisted…" Just after this: 
   "In the work, we developed *an* encoding for the 
    synthesis problem."

   Figure 1 legend, line 44: "typical" not "typic"

   Page 7, line 11: "We have tried to capture important 
   *aspects* of *the* Rothman-Schekman-Sudhof (RSS) model 
   of *the* vesicle…" Also "components" not "component"

   Example 1, line 40: "labels are a set of molecules"

   Example 2, line 36: "Fig. 3b presents its complete output VTS."

   Page 14, line 16: "We will also seek *to answer* two 
   kinds of questions, *given* the property."

   Page 15, line 14: "to ensure the k to be the least one" 
   -> "to ensure that k is minimized"

   Page 22: "The specification for *the* universal condition 
   *described in* Section 5"