### 2019-09-16 papers/journal/2019_ACTA_Rebuttal_Rd2.txt
Ashutosh rebuttal added

We would like to thank the reviewers for their very detailed and
helpful comments. In the following, we ellaborate on how we addressed
the feed in the revised version of our paper.

% Reviewer 1:
 > My comments and concerns have been addressed in detail. Thank you.

% Reviewer 3:

> I believe that most issues highlighted previously have been 
  addressed adequately. In particular, the structure of the 
  revised manuscript has been improved significantly. 
  I still found the paper challenging to read, but this is 
  largely due to the complexity of the biological system and 
  the necessary encoding details presented in the manuscript.

  Pt #1a:
-------- Mukund --------------
  > One point that I think requires further attention is the 
  motivation for using connectedness and the explanation of 
  this property, which is at the core of the proposed approach. 
  I did not find the additional explanations provided in the 
  response and the revised manuscript sufficient. Simply stating 
  that "The graph k-connectedness furnishes precisely" a test to 
  determine if information is missing from the model and that 
  "If the data […] does not have a certain degree of connectivity, 
  this implies that some biological data has been missed" does not 
  justify connectedness as a fundamental and required biological 
  property that all valid models should respect. The intuition of 
  connectedness as a robustness property provided elsewhere in the 
  manuscript helps but it is still not clear what evidence supports 
  the fact that VTSs are indeed robust. In addition, the problem as 
  formulated does not seem to seek the most robust networks as 
  explained in the following.

  Authors response: ...

  Pt #1b:
-------- Mukund [Subpoint added]--------------
  > Finally, it is not clear why the goal is to find the minimum graph 
    connectivity, given that larger connectivity corresponds to a 
    more robust network. I was expecting an approach involving 
    incrementing k to find the smallest value that makes the formula 
    unsatisfiable, thus identifying the maximum graph connectedness 
    possible for a given VTS with all other constraints being 
    satisfied. However, to the best of my understanding, this is not 
    what the paper is currently proposing, or else further  
    clarifications are needed.

  Authors response: ...

  > I also found the formulation of the main problems with respect 
    to connectedness somewhat confusing. For example, in Sec. 5.1.1 
    and Sec. 5.1.2 the problems are formulated as a search for VTS G 
    such that G is not k-connected. Yet, Sec. 5.1 specifies that 
    "In case the formula is satisfiable the procedure terminates and 
    we report the value of k", which seems contradictory.

  Authors response: In Sec. 5 we try to achieve two aims: give a 
  high-level explanation of the problem, followed by a formal search 
  statement (from a prespective of an efficient implementation). 
  This difference was only hinted but not explicitely stated. We have 
  now rewritten the parts of the text to address the distinction 
  between high-level overview and the implementation. 
  We have also removed the specific search problems (as suggested by 
  Reviewer #3 in the first revision) for further apparent accessibility 
  of the subject matter.

  > Another point that should be revised further is around 
    clarifying the contributions of this paper. The response 
    document summarizes nicely that, together with additional 
    QBF encoding, "this paper brings together all the results 
    with extended experiments and more detailed expositions for 
    the benefit of both biological and computation community." 
    However, I did not find a similar concise statement of the 
    contributions in the text. In addition, some of the stated 
    advances over previous approaches seem incremental and minor. 
    For example, for the previous CBMC-based method it is stated 
    that "...the scalability of the method was insufficient. 
    The tool was able to handle the system with just eight 
    compartments". Yet, we learn that "A typical eukaryotic 
    cell contains 9 organelles" and even one of the methods 
    presented in this paper fails for some of the experimental 
    evaluations with 7 or more compartments. The contributions 
    of this work, perhaps in the integration and evaluation of
    several methods reported previously, should be stated clearly 
    and accurately in the main text.

  Authors response: We clarified the concerns of the reviewers 
  and clearly stated in the Sec 1. Introduction that the CBMC 
  method does not scale and the direct encoding does. 
  We have also stated that the key purpose of this paper is 
  one unified exposition of our work.

  > A final minor point is about the use of "Synthesis" as 
    an encoding in Fig 4 and the relevant explanations. 
    Since the other problems list verification techniques 
    and solvers (e.g. BMC, SMT, QBF) and no novel synthesis 
    algorithms were developed for this problem, it would be 
    more appropriate to specify QBF as the encoding used for 
    the "VTS synthesis" problem.

  Authors response: We have now incorporated this information 
  in Fig 4.

% Reviewer 4:

  > Point 6 on page 3 is the same as point 4.

  Authors response: We thank the reviewer for pointing 
  this out. The two points were indeed duplicates. We 
  have now removed point 6.

  > I would argue still that the biological section has 
    unnecessary detail and could be simplified further.

  Authors response: We acknowledge that in the paper, 
  we only deal explicitly with SNARE proteins. Although 
  it is not required for an understanding of the work, 
  we feel a glimpse of the kinds of processes involved 
  in vesicle traffic, and its molecular machinery might 
  help the reader appreciate why we have built our model 
  in this particular way.

  > That the steady-state condition corresponds to the 
    biological meaning of homeostasis is repeated. This 
    only needs to be mentioned once.

  Authors response: We now state the equivalence of our 
  steady state condition and homeostasis of compartments
  only in section 2.3.2.

  > There is overlap in the 3rd and 6th paragraphs of 
    the 'Related Work' section, which refer to the use 
    of SAT/SMT in GRNs.

  Authors response: The use of SAT/SMT in GRN is presented
  in two different context (non-overlapping references). 
  The third paragrah points to general use of SAT/SMT 
  solvers in GRN, whereas 6th paragraph refers to a specific
  use (synthesis). We have updated the text to improve this
  distinction.

Minor points:

  > Small typos / grammatical errors. There are a number 
    of these throughout the text, some of which are listed 
    below (this is not exhaustive, so I recommend carefully 
    reading the manuscript to ensure clarity):
  Authors response: We thank the reviewer for pointing these 
  out. We have corrected the listed errors and performed 
  careful readings fixing other errors that we found.

  > Page 2, lines 11-12: Understanding how membrane traffic 
    functions *is*, therefore,…
  Authors response: Correction incorporated.

  > Page 2, line 13: Cell biologists have made… (not 'The')
  Authors response: Correction incorporated.

  > Page 3, line 14: "…parts of the traffic system that 
    *are* unchanging…"
  Authors response: Correction incorporated.

  > Page 4, line 6: "We also worked on the synthesis of the 
    unknown pieces *that* may be assisted…" Just after this: 
    "In the work, we developed *an* encoding for the 
    synthesis problem."
  Authors response: Both corrections are incorporated.

  > Figure 1 legend, line 44: "typical" not "typic"
  Authors response: Corrections incorporated 
  (typic was in caption of Fig. 1).

  > Page 7, line 11: "We have tried to capture important 
    *aspects* of *the* Rothman-Schekman-Sudhof (RSS) model 
    of *the* vesicle…" Also "components" not "component" 
  Authors response: Both corrections are incorporated.

  > Example 1, line 40: "labels are a set of molecules"
  Authors response: Correction incorporated.

  > Example 2, line 36: "Fig. 3b presents its complete output VTS."
  Authors response: Correction incorporated.

  > Page 14, line 16: "We will also seek *to answer* two 
    kinds of questions, *given* the property."
  Authors response: Both corrections are incorporated.

  > Page 15, line 14: "to ensure the k to be the least one" 
    -> "to ensure that k is minimized"
  Authors response: Correction incorporated.

  > Page 22: "The specification for *the* universal condition 
   *described in* Section 5" 
  Authors response: Both corrections are incorporated.