### 2019-09-22 papers/journal/2019_ACTA_Rebuttal_Rd2.txt
Mukund's push

We would like to thank the reviewers for their very detailed and helpful comments. In the following, we elaborate on how we addressed the feedback in the revised version of our paper.

% Reviewer 1:
  > My comments and concerns have been addressed in detail. 
    Thank you.
  Authors response: Thank you for your valuable comments.

% Reviewer 3:

> I believe that most issues highlighted previously have been 
  addressed adequately. In particular, the structure of the 
  revised manuscript has been improved significantly. 
  I still found the paper challenging to read, but this is 
  largely due to the complexity of the biological system and 
  the necessary encoding details presented in the manuscript.

  Authors response: Thank you for this feedback. As we are trying to reach an interdisciplinary audience we are forced to balance to some extent the biological detail with the technical discussions. We appreciate the feedback from the reviewers, which has helped us improve readability.

  > One point that I think requires further attention is the 
  motivation for using connectedness and the explanation of 
  this property, which is at the core of the proposed approach. 
  I did not find the additional explanations provided in the 
  response and the revised manuscript sufficient. Simply stating 
  that "The graph k-connectedness furnishes precisely" a test to 
  determine if information is missing from the model.

Authors response:

We have removed the word ``precisely'', which the Reviewer has correctly pointed out is too strong. We have replaced this by the sentence ``It has been shown that graph $k$-connectedness furnishes one such test''.

>  "If the data […] does not have a certain degree of connectivity,  this implies that some biological data has been missed" does not  justify connectedness as a fundamental and required biological  property that all valid models should respect. The intuition of  connectedness as a robustness property provided elsewhere in the  manuscript helps but it is still not clear what evidence supports  the fact that VTSs are indeed robust. In addition, the problem as  formulated does not seem to seek the most robust networks as  explained in the following.

Author response: We have added several paragraphs to motivate our focus on k-edge-connectedness:

``The informativeness and usefulness of using $k$-connectedness is motivated by the following example. ... In summary, the higher the edge-connectedness of a measured VTS graph, the more of the underlying biological constraints can be satisfied by appropriately placing SNAREs on each node and edge.''

  > Finally, it is not clear why the goal is to find the minimum graph connectivity, given that larger connectivity corresponds to a more robust network. I was expecting an approach involving incrementing k to find the smallest value that makes the formula unsatisfiable, thus identifying the maximum graph connectedness possible for a given VTS with all other constraints being satisfied. However, to the best of my understanding, this is not what the paper is currently proposing, or else further clarifications are needed.

Author response: We apologise for the confusing structure of our text. To clarify this point we have added a new paragraph:

``It is perhaps counter-intuitive that, beyond a certain degree of edge-connectedness, any VTS graph can be implemented as a biological system. This is because we have not placed any constraints on the number of distinct SNAREs and other trafficking molecules a cell contains. In practice there is an upper bound on the number of molecules a cell can genetically encode. This question (``What is the maximum edge-connectedness possible given a certain number of SNAREs?'') is complementary to the question we have asked here (``What is the minimum edge-connectedness possible given any number of SNARES operating under certain physical constraints?'').''

  > I also found the formulation of the main problems with respect to connectedness somewhat confusing. For example, in Sec. 5.1.1 and Sec. 5.1.2 the problems are formulated as a search for VTS G such that G is not k-connected. Yet, Sec. 5.1 specifies that "In case the formula is satisfiable the procedure terminates and we report the value of k", which seems contradictory.

  Authors response: In Sec. 5 we try to give a high-level 
  explanation of the procedure, followed by a formal search statement from the perspective of an efficient implementation. 
  This difference was only hinted but not explicitly stated. We have  now rewritten the parts of the text to address the distinction between high-level overview and the implementation. We have also removed the specific search problems (as suggested by Reviewer #3 in the first revision) for further apparent accessibility of the subject matter.

  > Another point that should be revised further is around 
    clarifying the contributions of this paper. The response 
    document summarizes nicely that, together with additional 
    QBF encoding, "this paper brings together all the results 
    with extended experiments and more detailed expositions for 
    the benefit of both biological and computation community." 
    However, I did not find a similar concise statement of the 
    contributions in the text. In addition, some of the stated 
    advances over previous approaches seem incremental and minor. 
    For example, for the previous CBMC-based method it is stated 
    that "...the scalability of the method was insufficient. 
    The tool was able to handle the system with just eight 
    compartments". Yet, we learn that "A typical eukaryotic 
    cell contains 9 organelles" and even one of the methods 
    presented in this paper fails for some of the experimental 
    evaluations with 7 or more compartments. The contributions 
    of this work, perhaps in the integration and evaluation of
    several methods reported previously, should be stated clearly 
    and accurately in the main text.

  Authors response: We clarified the concerns of the reviewers 
  and clearly stated in the Sec 1. Introduction that the CBMC 
  method does not scale and the direct encoding does. 
  We have also stated that the key purpose of this paper is 
  one unified exposition of our work.

  > A final minor point is about the use of "Synthesis" as 
    an encoding in Fig 4 and the relevant explanations. 
    Since the other problems list verification techniques 
    and solvers (e.g. BMC, SMT, QBF) and no novel synthesis 
    algorithms were developed for this problem, it would be 
    more appropriate to specify QBF as the encoding used for 
    the "VTS synthesis" problem.

  Authors response: We have now incorporated this information 
  in Fig 4.

% Reviewer 4:

  > Point 6 on page 3 is the same as point 4.

  Authors response: We thank the reviewer for pointing 
  this out. The two points were indeed duplicates. We 
  have now removed point 6.

  > I would argue still that the biological section has 
    unnecessary detail and could be simplified further.

  Authors response: We acknowledge that in the paper, 
  we only deal explicitly with SNARE proteins. Although 
  it is not required for an understanding of the work, 
  we feel a glimpse of the kinds of processes involved 
  in vesicle traffic, and its molecular machinery might 
  help the reader appreciate why we have built our model 
  in this particular way.

  > That the steady-state condition corresponds to the 
    biological meaning of homeostasis is repeated. This 
    only needs to be mentioned once.

  Authors response: We now state the equivalence of our 
  steady state condition and homeostasis of compartments
  only in section 2.3.2.

  > There is overlap in the 3rd and 6th paragraphs of 
    the 'Related Work' section, which refer to the use 
    of SAT/SMT in GRNs.

  Authors response: The use of SAT/SMT in GRN is presented
  in two different context (non-overlapping references). 
  The third paragrah points to general use of SAT/SMT 
  solvers in GRN, whereas 6th paragraph refers to a specific
  use (synthesis). We have updated the text to improve this
  distinction.

Minor points:

  > Small typos / grammatical errors. There are a number 
    of these throughout the text, some of which are listed 
    below (this is not exhaustive, so I recommend carefully 
    reading the manuscript to ensure clarity):
  Authors response: We thank the reviewer for pointing these 
  out. We have corrected the listed errors and performed 
  careful readings fixing other errors that we found.

  > Page 2, lines 11-12: Understanding how membrane traffic 
    functions *is*, therefore,…
  Authors response: Correction incorporated.

  > Page 2, line 13: Cell biologists have made… (not 'The')
  Authors response: Correction incorporated.

  > Page 3, line 14: "…parts of the traffic system that 
    *are* unchanging…"
  Authors response: Correction incorporated.

  > Page 4, line 6: "We also worked on the synthesis of the 
    unknown pieces *that* may be assisted…" Just after this: 
    "In the work, we developed *an* encoding for the 
    synthesis problem."
  Authors response: Both corrections are incorporated.

  > Figure 1 legend, line 44: "typical" not "typic"
  Authors response: Corrections incorporated 
  (typic was in caption of Fig. 1).

  > Page 7, line 11: "We have tried to capture important 
    *aspects* of *the* Rothman-Schekman-Sudhof (RSS) model 
    of *the* vesicle…" Also "components" not "component" 
  Authors response: Both corrections are incorporated.

  > Example 1, line 40: "labels are a set of molecules"
  Authors response: Correction incorporated.

  > Example 2, line 36: "Fig. 3b presents its complete output VTS."
  Authors response: Correction incorporated.

  > Page 14, line 16: "We will also seek *to answer* two 
    kinds of questions, *given* the property."
  Authors response: Both corrections are incorporated.

  > Page 15, line 14: "to ensure the k to be the least one" 
    -> "to ensure that k is minimized"
  Authors response: Correction incorporated.

  > Page 22: "The specification for *the* universal condition 
   *described in* Section 5" 
  Authors response: Both corrections are incorporated.